Prior answering the questions of the assignment, there are few steps are required to be executed as a preparation which will enable me to answer the questions.

I will first start with uploading the relevant libraries that we might use during developing the relevant code for the analysis.

```{r}
library(dplyr) ; library(knitr) ; library(ggplot2)
```

Thereafter, I will read the data and stored it into a variable, and do the needed processing to be ready for the analysis.

```{r, echo=TRUE}
file1<-read.csv("activity.csv",header=TRUE) %>% tbl_df
file<-filter(file1,!is.na(steps))
```

Now, we should be ready for the analysis part and answering the questions.

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r, echo=TRUE}
# summation of steps per day
file.sum.steps.day<-group_by(file,date) %>% summarize(sum(steps))
# print the 1st 10 row that shows the total number of steps taken per day.
head(file.sum.steps.day,10)
```


2. Make a histogram of the total number of steps taken each day

```{r, echo=TRUE}
# histogram
hist(file.sum.steps.day$"sum(steps)",nclass=6, main="Histogram of the total number of steps per day", xlab="Total no. of steps")
```

3. Calculate and report the mean and median of the total number of steps taken per day

```{r, echo=TRUE}
file.mean.sum.steps.day<-summary(file.sum.steps.day[2])[4]
file.median.sum.steps.day<-summary(file.sum.steps.day[2])[3]
```

Thus the mean of the total no. of steps is `r file.mean.sum.steps.day` and the `r file.median.sum.steps.day`


## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
# Calculate the average 
steps.interval<-aggregate(steps~interval,file,mean)
# Plot the outcome
plot(steps.interval$interval,steps.interval$steps,type="l",xlab="Interval",ylab="Avg.no. of Steps", main="Avg of steps accross all days")
```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE}
max.steps<-steps.interval[which.max(steps.interval$steps),]
max.steps
```

The interval of `r max.steps[1]` has the maximum average value of steps `r max.steps[2]`.

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, echo=TRUE}
total.no.missing.data<-nrow(file1)-nrow(file)
total.no.missing.data
```

The total no. of "NA" is `r total.no.missing.data`.


2. Devise a strategy for filling in all of the missing values in the dataset. I have filled the NA field with the mean of the steps. 

```{r, echo=TRUE}
file1[is.na(file1[1]),][1]<-mean(file$steps)
```


3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo=TRUE}
head(file1,10)
```


4.a Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r, echo=TRUE}
file1.sum.steps.day<-group_by(file1,date) %>% summarize(sum(steps))
# histogram
hist(file1.sum.steps.day$"sum(steps)",nclass=6,main="Histogram of total number of steps per day with replaced NA with the mean", xlab="Total no. of steps per day")
```


4.b Do these values differ from the estimates from the first part of the assignment? 

```{r, echo=TRUE}
file1.mean.sum.steps.day<-summary(file1.sum.steps.day[2])[4]
file1.median.sum.steps.day<-summary(file1.sum.steps.day[2])[3]

file1.mean.sum.steps.day ; file1.median.sum.steps.day
```

As we can see that there is no significant changes from imputing the missing data NA. as the mean value of the steps with and without the missing data is `r file.mean.sum.steps.day` and `r file1.mean.sum.steps.day` respectively. The median value of the steps with and without the missing data is `r file.median.sum.steps.day` and `r file1.median.sum.steps.day` respectively. Thus, the mean values are the same however, there is a light difference in the median.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day

```{r, echo=TRUE}
# Converting the date into week day
file1<- mutate(file1,weekday=weekdays(as.Date(date)))

# Splittin the weekday into either weekend or weekday
file1$weekday[file1$weekday %in% c("Saturday","Sunday")]<-"weekend"
file1$weekday[file1$weekday !="weekend"]<-"weekday"
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r, echo=TRUE}
file.mean.steps.per.day<-aggregate(steps ~ interval + weekday, file1, mean)
qplot(interval,steps,data=file.mean.steps.per.day,type="l",geom=c("line"),xlab="Interval", ylab="Number of steps")+facet_wrap(~ weekday,ncol=1)
```


